---
title: "Case study_Predictiton"
output: html_notebook
---

```{r}
library(dummies)
library(ggplot2)
library(car)
library(kader)
library(MASS)
library(scales)
library(psych)
library(e1071)
library(outliers)
library(rpart)
library(rpart.plot)
library(randomForest)
library(caret)
library(mgcv)
library(ROCR)
library(gains)
library(Metrics)
library(caTools)
library(e1071)
library(gmodels)
library(corrplot)
library(hydroGOF)
library(psycho)
library(gridExtra)
library(grid)
library(pROC)

```


```{r Import cleaned data}
Music_ <- read.csv(choose.files())
str(Music_)

```

```{r Subsets for prediction}

Music1 <- Music_[,c(2:13)]
Music1.notation <- dummy(Music1$notation, sep = "_")
Music <- cbind(Music1[,c(1:2, 4:11)], Music1.notation[,-12])
#music without notation
Music2 <- Music1[,-4]
#Music for classification
Music.cl <- Music_[,c(3:4, 6:14)]

#use PCA to do dimension reduction
fa.parallel(Music[,-1], fa="pc", n.iter = 100,show.legend = FALSE, main = "Scree plot")#choose 5 pcs
pc <- principal(Music[,-1], nfactors = 5, rotate="promax",
                scores = TRUE)
head(pc$scores)
round(pc$weights,2)

#data tranformation for normality
#acousticness
qplot(x=acousticness, bins=30, data = Music)#very skewed
qplot(kader:::cuberoot(Music$acousticness))
Music$acousticness <- -log((1/(kader:::cuberoot(Music$acousticness)))-1, base = exp(1))
qplot(log(Music$acousticness))#best tranformation
{qqnorm(Music$acousticness)
qqline(Music$acousticness)}
qqPlot(Music$acousticness)

#Danceability
qplot(Music$danceability)#left skewed
#log trans
qplot(log(Music$danceability, base = exp(1)))
qplot(log1p(Music$danceability))
qplot(1/Music$danceability)
qplot(sqrt(Music$danceability))
qplot(kader:::cuberoot(Music$danceability))
qplot(log1p(kader:::cuberoot(Music$danceability)))

#normalize new predictors
normalize <- function(x){
  return((x-min(x))/(max(x)-min(x)))
}
rc1.nor <- normalize(pc$scores[,1])
rc2.nor <- normalize(pc$scores[,2])
rc3.nor <- normalize(pc$scores[,3])
rc4.nor <- normalize(pc$scores[,4])
rc5.nor <- normalize(pc$scores[,5])

RC <- data.frame(rc1.nor, rc2.nor, rc3.nor, rc4.nor, rc5.nor) 
#RC1 transformation
rc1 <- RC[,1]
{hist(rc1, prob=TRUE)
lines(density(rc1))}
{hist(rc1^5, prob=TRUE)
lines(density(rc1^5))}
engine.displ <- skewness(rc1^6)
engine.displ #0.112 onlu moderately skewed
outlier(rc1^6)
RC$rc1.nor <- (RC$rc1.nor)^6

#RC2 transforamtion
rc2 <- RC[,2]
{hist(rc2, prob=TRUE)
lines(density(rc2))}
engine.displ <- skewness(rc2)
engine.displ

#RC3 transformation
rc3 <- RC[,3]
{hist(rc3, prob=TRUE)
lines(density(rc3))}
{hist(rc3^2, prob=TRUE)
lines(density(rc3^2))}
engine.displ <- skewness(rc3^2)
engine.displ
RC$rc3.nor <- (RC$rc3.nor)^2

#RC4 transformation
rc4 <- RC[,4]
{hist(rc4, prob=TRUE)
lines(density(rc4))}
{hist(rc4^2, prob=TRUE)
lines(density(rc4^2))}
engine.displ <- skewness(rc4^2)
engine.displ
RC$rc4.nor <- (RC$rc4.nor)^2 

#RC5 transformation
rc5 <- RC[,5]
{hist(rc5, prob=TRUE)
lines(density(rc5))}
engine.displ <- skewness(rc5)
engine.displ

RC <- cbind(Music$popularity, RC)
colnames(RC) <- c("Pop", "RC1", "RC2", "RC3", "RC4", "RC5")

```

```{r partition the datasets and visualize the relationships}

set.seed(123)
#partitionin by using RC
train.index1 <- sample(1:8986, 0.5*dim(RC[1]))
RCTrain <- RC[train.index1,]
RCV_T <- RC[-train.index1,]

Valid.index1 <- sample(1:3595, 0.6*dim(RCV_T[1]))
RCValid <- RCV_T[Valid.index1,]
RCTest <- RCV_T[-Valid.index1,]

#partitioning by using Music1
set.seed(123)
colnames(Music) <- c( "popularity","key","acousticness","danceability",      "duration_minutes","liveness","loudness_decibels","speechiness","tempo_bpm","valence", "notation_A","notation_A#_Bb","notation_B","notation_C","notation_C#_Db","notation_D","notation_D#_Eb","notation_E","notation_F","notation_F#_Gb","notation_G")
train.index2 <- sample(1:8986, 0.5*dim(Music1[1]))
MusicTrain <- Music1[train.index2,]
MusicV_T <- Music1[-train.index2,]

Valid.index2 <- sample(1:3595, 0.6*dim(MusicV_T[1]))
MusicValid <- MusicV_T[Valid.index2,]
MusicTest <- MusicV_T[-Valid.index2,]

#partioning by music
set.seed(123)
train.index3 <- sample(1:8986, 0.5*dim(Music[1]))
MusicTrain1 <- Music[train.index3,]
MusicV_T1 <- Music[-train.index3,]

Valid.index3 <- sample(1:3595, 0.6*dim(MusicV_T1[1]))
MusicValid1 <- MusicV_T1[Valid.index3,]
MusicTest1 <- MusicV_T1[-Valid.index3,]

#partionning by music2(without notation)
set.seed(100)
train.index4 <- sample(1:8986, 0.5*dim(Music2[1]))
MusicTrain2 <- Music2[train.index4,]
MusicV_T2 <- Music2[-train.index4,]

Valid.index4 <- sample(1:3595, 0.6*dim(MusicV_T2[1]))
MusicValid2 <- MusicV_T2[Valid.index4,]
MusicTest2 <- MusicV_T2[-Valid.index4,]

#partionning by music.cl
set.seed(123)
train.index5 <- sample(1:8986, 0.5*dim(Music.cl[1]))
MusicTrain.cl <- Music.cl[train.index5,]
MusicV_T.cl <- Music.cl[-train.index5,]

Valid.index5 <- sample(1:3595, 0.6*dim(MusicV_T.cl[1]))
MusicValid.cl <- MusicV_T.cl[Valid.index5,]
MusicTest.cl <- MusicV_T.cl[-Valid.index5,]

#visualize 
attach(RCTrain)
scatter.smooth(RC1,Pop)
scatter.smooth(RC2,Pop)
scatter.smooth(RC3,Pop)
scatter.smooth(RC4,Pop)
scatter.smooth(RC5,Pop)
```

```{r multiple linear regression}
#fit regression models with lm()

#model 1
cor(RCTrain)
music.fit1 <- lm(Pop ~ RC1+RC2+RC3+RC4+RC5, data = RCTrain)
summary(music.fit1)

#regression diagnostic
confint(music.fit1)
##normality: enhanced approach to evaluate
qqPlot(music.fit1, labels=row.names(RCTrain), id.method="identify", simulate=TRUE, main="Q-Q Plot")

#This model violates the normality
```

```{r Random forest }
rf <- randomForest(popularity~., data = MusicTrain2)
rf
importance(rf)
rf.pred.train <- predict(rf, MusicTrain2)
rf.pred.valid <- predict(rf, MusicValid2)
rf.pred.test <- predict(rf, MusicTest2)

#use MAE
MAE(rf.pred.train, MusicTrain2$popularity) #2.608
MAE(rf.pred.valid, MusicValid2$popularity)#5.925
MAE(rf.pred.test, MusicTest2$popularity) #6.728

#MAPE
mape(rf.pred.train, MusicTrain2$popularity) #0.047
mape(rf.pred.valid, MusicValid2$popularity)#0.106
mape(rf.pred.test, MusicTest2$popularity) #.121

#use rmse to evaluate the performance
rmse.train <- RMSE(rf.pred.train, MusicTrain2$popularity)#3.326
rmse.valid <- RMSE(rf.pred.valid, MusicValid2$popularity)#7.746
rmse.test <- RMSE(rf.prerd.test, MusicTest2$popularity)#8.031

#Use Mean error
me(rf.pred.train, MusicTrain2$popularity)#0.077
me(rf.pred.valid, MusicValid2$popularity)#-1.433
me(rf.pred.test, MusicTest2$popularity)#2.661

gain <- gains(MusicValid2$popularity,
              rf.pred.valid, groups = 1000)
{plot(c(0, gain$cume.pct.of.total*sum(MusicValid2$popularity))~c(0, gain$cume.obs))
lines(c(0, sum(MusicValid2))~c(0, dim(MusicValid2)[1]))}

#The RMSE shows that this model is good.

```

```{r SVM}
mu.svm <- svm(Music2[,-1], Music2[,1])
data.frame(actual=Music2[,1], predicted=mu.svm$fitted)
rmse.svm <- RMSE(mu.svm$fitted, Music2$popularity)#7.52
me(mu.svm$fitted, Music2$popularity)
MAE(mu.svm$fitted, Music2$popularity)
mape(mu.svm$fitted, Music2$popularity)

lift.svm <- lift(relevel(as.factor(Music.cl$popularcategory), ref="1") ~ mu.svm$fitted)
xyplot(lift.svm, plot = "gain")


#this model is good

```

```{r LDA}

mu.lda <- lda(popularcategory~., MusicTrain.cl)
mu.lda.pred <- predict(mu.lda, as.data.frame(MusicValid.cl[,-11]))
data.frame(actual=MusicValid.cl[,11], predicted=mu.lda.pred$class)
confusionMatrix(as.factor(MusicValid.cl$popularcategory), as.factor(mu.lda.pred$class))

lift.lda <- lift(relevel(as.factor(MusicValid.cl$popularcategory), ref="1") ~ mu.lda.pred$class)
xyplot(lift.lda, plot = "gain")

r <- roc(as.numeric(MusicValid.cl$popularcategory), as.numeric(mu.lda.pred$class))
plot.roc(r)

auc(r)
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
